# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("~/repos/RepData_PeerAssessment1/")
file <- unzip("activity.zip")
data <- read.csv(file)
data$date <- as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r}
data.total <- tapply(data$steps, data$date, sum)
hist(data.total, main="Total number of steps taken each day", xlab="")
```

```{r resluts="asis"}
mean(data.total, na.rm=T)
median(data.total, na.rm=T)
```

## What is the average daily activity pattern?
```{r}
library(reshape2)
data.long <- melt(data, id.vars = c("interval", "date"))
data.activity <- dcast(data.long, interval ~ variable, mean, na.rm=T)
plot(data.activity$interval, data.activity$steps, type="l", xlab="5-min interval", ylab="Average number of steps", main="Averaged number of steps per interval across all days")
```

Interval with the maximum number of steps
```{r, results="asis"}
data.activity$interval[which.max(data.activity$steps)]
```

## Imputing missing values
Number of NA samples in the dataset
```{r, results="asis"}
sum(!complete.cases(data))
```

Let's impute NA values with 0
```{r}
dataset <- data
dataset$steps[is.na(dataset$steps)] <- 0
dataset.total <- tapply(dataset$steps, dataset$date, sum)
```

Number of steps taken each day after NA values were imputed
```{r}
hist(dataset.total, main="Total number of steps taken each day", xlab="")
```

```{r resluts="asis"}
mean(dataset.total)
median(dataset.total)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(reshape2)
library(ggplot2)
data$days <- weekdays(data$date)
data$days <- gsub("Sunday|Saturday", "Weekend", data$days)
data$days <- gsub("Monday|Tuesday|Wednesday|Thursday|Friday", "Weekday", data$days)
data$days <- as.factor(data$days)
data.long <- melt(data[,c("interval","steps","days")], id.vars = c("interval", "days"))
data.activity <- dcast(data.long, interval + days~ variable, mean, na.rm=T)
qplot(interval, steps, data=data.activity, facets=days~., geom="line", main="Averaged number of steps per interval across all weekdays")
```
